Observations from Parallelising Three Maximum Common (Connected) Subgraph Algorithms

### 2017-08-09 paper/paper.tex
This is definitely progress

@@

%Conference
\acmConference[$\textnormal{IA}^3$ 2017]{Seventh Workshop on Irregular Applications: Architectures and Algorithms}{November 2017}{Denver, Colorado USA}
\acmYear{2017}
\copyrightyear{2017}

@@

Finding a maximum common subgraph is the key step in measuring the similarity or difference between
two graphs \citep{DBLP:journals/prl/Bunke97,DBLP:journals/prl/FernandezV01,o:Kriege15}: to determine
the difference between two graphs, we find what they have in common, and then take everything left
over. Because of this, maximum common subgraph problems frequently arise in biology and chemistry
\citep{DBLP:journals/jcamd/RaymondW02a,o:EhrlichR11,DBLP:journals/dam/GayFMSS14} where graphs
represent molecules, but also in applications including computer vision
\citep{DBLP:journals/jair/CookH94,DBLP:conf/gbrpr/CombierDS13}, computer-aided manufacturing
\citep{o:LuoWSN17}, crisis management \citep{o:DelavalladeFLL16}, deanonymising datasets
\citep{o:SharadD13}, the analysis of source code \citep{DBLP:journals/tkde/DjokoCH97}, binary
programs \citep{DBLP:conf/icics/GaoRS08}, and circuit designs \citep{DBLP:journals/jair/CookH94}, in
malware detection \citep{DBLP:journals/compsec/ParkRS13}, social network analysis
\citep{DBLP:journals/tkde/FangYZZ15}, graph database query explanations
\citep{DBLP:journals/jcss/VasilyevaTBL16}, and in character recognition problems
\citep{DBLP:journals/pr/LuRS91}.  A common variant of the problem requires a largest
\emph{connected} subgraph
\citep{DBLP:journals/tcs/Koch01,DBLP:journals/jcamd/RaymondW02a,DBLP:conf/mco/VismaraV08,o:EhrlichR11,o:LuoWSN17}.
As illustrated in \cref{figure:mcsexample}, the maximum common connected subgraph cannot be deduced
from the maximum common subgraph.

@@

\subsection{Encoding as Maximum Clique}

@@

The maximum common subgraph problem may be reformulated as a constraint optimisation problem, as
follows.  Observe that an equivalent definition of a common subgraph of graphs $G$ and $H$ is an
injective partial mapping from $G$ to $H$ which preserves both adjacency and non-adjacency. Hence we
pick whichever input graph has fewer vertices, and call it the \emph{pattern}; the other graph is
called the \emph{target}. The model is then to construct an injective, partial mapping from the
pattern to the target, preserving both adjacency and non-adjacency. Thus, for each vertex in the
pattern, we create a variable, whose domain ranges over each vertex in the target graph, plus an
additional value $\bot$. We then have three sets of constraints. The first set says that for each
pair of adjacent vertices in the pattern (that is, for each edge in the pattern), if neither of
these vertices are mapped to $\bot$ then these vertices must be mapped to an adjacent pair of target
vertices. The second set is similar, but looks at non-adjacent pairs (or non-edges).  Finally, the
third set ensures injectivity, by enforcing that the variables must be all different except when
using $\bot$.

?? Either by binary constraints between all pairs, or a specialised propagator.

?? Objective.

?? One approach is branch and bound search over this model. Pick a variable, ... . cite.

@@

Due to the special structure of the maximum common subgraph problem, the following property holds
throughout the search process using the CP model: any two variables either have domains with no
values in common (with the possible exception of $\bot$), or have identical domains. The algorithm
McSplit exploits this property, while exploring essentially the same search tree as the CP model.
Rather than storing a domain for each vertex in $\operatorname{V}(G)$, which results in redundant
storage in cases where domains are identical, equivalence classes of vertices in
$\operatorname{V}(G)$ are stored, and a set of candidate vertices in $H$ is maintained for each
equivalence class. This enables fast propagation of the constraints and smaller memory requirements.
In addition, the data structure enables good branching heuristics to be calculated cheaply.

@@

The algorithm as described by Hoffmann, McCreesh, and Reilly \cite{DBLP:conf/aaai/HoffmannMR17} approaches the subgraph isomorphism problem by asking the question ``if a pattern graph cannot be found in the target, how much of the pattern graph can be found?''. The $k{\downarrow}$ algorithm tries to solve the subgraph isomorphism problem first for $k=0$ (the whole pattern graph can be found in the target), should that be not satisfiable, it tries to solve the problem for $k=1$ (one vertex cannot be matched) should that also not be satisfiable, it iterates $k$ further by a small number to keep the shrinking pattern graph within a reasonable size. By unmatching some vertices  invariants are weakened but kept effective. Specifically, invariants based around paths and degrees of vertices hold.

This technique bridges the subgraph isomorphism problem with the maximal common subgraph problem. The iteration of the decision problem (iterating over $k$) leads to an algorithm with rivalling runtimes.

@@

?? Approach

@@

>>>>

\usepackage{placeins}

@@

%Conference
\acmConference[IA$^3$ 2017]{Seventh Workshop on Irregular Applications: Architectures and Algorithms}{November 2017}{Denver, Colorado USA}
\acmYear{2017}
\copyrightyear{2017}

@@

Finding a maximum common subgraph is the key step in measuring the similarity or difference between
two graphs \citep{DBLP:journals/prl/Bunke97,DBLP:journals/prl/FernandezV01,o:Kriege15}: to determine
the difference between two graphs, we find what they have in common, and then take everything left
over. Because of this, maximum common subgraph problems frequently arise in biology and chemistry
\citep{DBLP:journals/jcamd/RaymondW02a,o:EhrlichR11,DBLP:journals/dam/GayFMSS14} where graphs
represent molecules, but also in applications including computer vision
\citep{DBLP:journals/jair/CookH94,DBLP:conf/gbrpr/CombierDS13}, computer-aided manufacturing
\citep{o:LuoWSN17}, crisis management \citep{o:DelavalladeFLL16}, deanonymising datasets
\citep{o:SharadD13}, the analysis of source code \citep{DBLP:journals/tkde/DjokoCH97} and binary
programs \citep{DBLP:conf/icics/GaoRS08}, in
malware detection \citep{DBLP:journals/compsec/ParkRS13}, social network analysis
\citep{DBLP:journals/tkde/FangYZZ15}, graph database query explanations
\citep{DBLP:journals/jcss/VasilyevaTBL16}, and in character recognition problems
\citep{DBLP:journals/pr/LuRS91}.  A common variant of the problem requires a largest
\emph{connected} subgraph
\citep{DBLP:journals/tcs/Koch01,DBLP:journals/jcamd/RaymondW02a,DBLP:conf/mco/VismaraV08,o:EhrlichR11,o:LuoWSN17}.
As illustrated in \cref{figure:mcsexample}, a maximum common connected subgraph cannot be determined
from the maximum common subgraph.

@@

There are three competitive approaches for the maximum common subgraph problem, each being the
strongest on certain classes of instance. The first involves a reduction to the maximum clique
problem, whilst the other two approaches have a constraint programming feel to them.

\subsection{Reduction to Maximum Clique}

A clique in a graph is a subgraph where every vertex is adjacent to every other. There is a
well-known reduction from the maximum common subgraph problem to the problem of finding a maximum
clique in an \emph{association graph}
\citep{o:Levi73,DBLP:journals/siamcomp/BalasY86,o:DurandPBT99,DBLP:journals/jcamd/RaymondW02a,DBLP:conf/cp/McCreeshNPS16}, which, when combined with
a modern maximum clique solver \citep{DBLP:journals/ol/SegundoMRH13}, is the current best approach for
solving the problem on labelled graphs \citep{DBLP:conf/cp/McCreeshNPS16}. A modified clique-like
algorithm can also be used to solve the maximum common connected subgraph problem, by ensuring
connectedness during search \citep{DBLP:conf/cp/McCreeshNPS16}; again, this is the best known way of
solving the problem on labelled graphs.

A major downside of the maximum clique approach is that the association graph encoding is extremely
memory-intensive, limiting its practical use to pairs of graphs with no more than a few hundred
vertices.

@@

The maximum common subgraph problem may be reformulated as a constraint optimisation problem, as
follows.  Observe that an equivalent definition of a common subgraph of graphs $G$ and $H$ is an
injective \emph{partial} mapping from $G$ to $H$ which preserves both adjacency and non-adjacency.
Hence we pick whichever input graph has fewer vertices, and call it the \emph{pattern}; the other
graph is called the \emph{target}. The model then follows from this new definition: for each vertex
in the pattern, we create a variable, whose domain ranges over each vertex in the target graph, plus
an additional value $\bot$ representing an unmapped vertex. We then have three sets of constraints.
The first set says that for each pair of adjacent vertices in the pattern (that is, for each edge in
the pattern), if neither of these vertices are mapped to $\bot$ then these vertices must be mapped
to an adjacent pair of target vertices. The second set is similar, but looks at non-adjacent pairs
(or non-edges).  Finally, the third set ensures injectivity, by enforcing that the variables must be
all different except when using $\bot$. This final set of constraints may either be implemented
using binary constraints between all pairs of variables, or a special global ``all different except
$\bot$'' propagator \citep{DBLP:conf/cp/PetitRB01}. The objective is simply to find an assignment of
values to variables, maximising the number of variables not set to $\bot$. The state of the art for
this technique is a dedicated implementation of a forward-checking branch and bound search over this
model \citep{DBLP:conf/cp/NdiayeS11,DBLP:conf/cp/McCreeshNPS16}.

@@

\citet{o:McCreeshPT17} observe that due to the special structure of the maximum common subgraph
problem, the following property holds throughout the search process using the constraint programming
model: any two variables either have domains with no values in common (with the possible exception
of $\bot$), or have identical domains. The McSplit algorithm exploits this property. It explores
essentially the same search tree as the basic forward-checking constraint programming model, but
using different supporting algorithms and data structures.  Rather than storing a domain for each
vertex in the pattern graph, equivalence classes of vertices in both graphs are stored in a special
data structure which is modified in-place and restored upon backtracking. This enables fast
propagation of the constraints and smaller memory requirements. In addition, this data structure
enables stronger branching heuristics to be calculated cheaply.

@@

A different take on the constraint programming model is presented by
\citet{DBLP:conf/aaai/HoffmannMR17}. They approach the maximum common subgraph via the
subgraph isomorphism problem, asking the question ``if a pattern graph cannot be found in the
target, how much of the pattern graph can be found?''. The $k{\downarrow}$ algorithm tries to solve
the subgraph isomorphism problem first for $k=0$ (asking whether the whole pattern graph can be found in the
target). Should that be not satisfiable, it tries to solve the problem for $k=1$ (one vertex cannot
be matched), and should that also not be satisfiable, it iteratively increases $k$ until the result
is satisfiable. The advantage of this approach is that strong invariants using paths and the degrees
of vertices may be used to prune large portions of the search space.

This algorithm is aimed primarily at large instances, where the two graphs are of different orders,
and where it is expected that the solution will involve most of the smaller graph (that is, $k$ is
expected to be low). The implementation we work with only handles unlabelled instances, and does
not support the connected variant.

@@

\subsection{Anomalies}

\subsection{Parallel Maximum Clique}

?? For the clique algorithm.

?? Discuss the TOPC paper.

?? Reproducible

\subsection{Parallel Constraint-Based Search}

?? Approach for kdown and mcsplit

@@

?? There is more to be done.

\FloatBarrier
